---
title: "Bootstrap standard errors for TSPS and TSRI estimators"
author: "TP"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
---

```{r include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

```{r setup, message=FALSE}
library(OneSampleMR)
library(boot)
library(tidyverse)
```

## Simulate some data

```{r}
set.seed(9)
n <- 1000
psi0 <- 0.5
Z <- rbinom(n, 1, 0.5)
X <- rbinom(n, 1, 0.7 * Z + 0.2 * (1 - Z))
m0 <- plogis(1 + 0.8 * X - 0.39 * Z)
Y <- rbinom(n, 1, plogis(psi0 * X + log(m0 / (1 - m0))))
dat <- data.frame(Z, X, Y)
```

## Bootstrapping

* For bootstrapping first define a function which returns the estimate, first for logistic TSPS

```{r}
tspsest <- function(data, indices){
  d <- data[indices, ]
  stage1 <- lm(X ~ Z, data = d)
  d$xhat <- fitted.values(stage1)
  stage2 <- glm(Y ~ xhat, family = binomial, data = d)
  coef(stage2)[2]
}
```

* Test this returns the causal log OR
```{r}
tspsest(dat, 1:1000)
```

* Perform bootstrapping

```{r}
set.seed(123456)
tspsbt <- boot(dat, tspsest, R = 250)
summary(tspsbt)
boot.ci(tspsbt, type = "basic")
```

* For logistic TSRI

```{r}
tsriest <- function(data, indices){
  d <- data[indices, ]
  stage1 <- lm(X ~ Z, data = d)
  d$res <- residuals(stage1)
  stage2 <- glm(Y ~ X + res, family = binomial, data = d)
  coef(stage2)[2]
}
```

* Test this returns the causal log OR

```{r}
tsriest(dat, 1:1000)
```

* Perform bootstrapping

```{r}
set.seed(7891011)
tsribt <- boot(dat, tsriest, R = 250)
summary(tsribt)
boot.ci(tsribt, type = "basic")
```

## Comparison estimates and SEs using GMM from the OneSampleMR package

* Example logistic TSPS estimate with SEs using GMM estimation in my **OneSampleMR** package

```{r}
tspslogitfit <- tsps(Y ~ X | Z , data = dat, link = "logit")
summary(tspslogitfit)
```

* Example logistic TSRI estimate with SEs using GMM estimation in my **OneSampleMR** package

```{r}
tsrilogitfit <- tsri(Y ~ X | Z , data = dat, link = "logit")
summary(tsrilogitfit)
```

## Manual estimation with unadjusted SEs

* Stage 1

```{r}
stage1 <- lm(X ~ Z, data = dat)
dat$xhat <- fitted.values(stage1)
dat$res <- residuals(stage1)
```

* Logistic TSPS

```{r}
tspss2 <- glm(Y ~ xhat, family = binomial, data = dat)
summary(tspss2)
```

* Logistic TSRI

```{r}
tsris2 <- glm(Y ~ X + res, family = binomial, data = dat)
summary(tsris2)
```

## R session information

<details>
<summary>For reproducibility</summary>
```{r}
quarto::quarto_version()
sessioninfo::session_info()
```
</details>
